---
title: "Walkthrough 2: Approaching Gradebook Data From a Data Science Perspective"
output: html_notebook
---
This walkthrough further explores the ubiquitous but not-often-analyzed classroom gradebook dataset. We will use data science tools and techniques and focus more on analyses, including correlations and linear models.
```{r}
# Load libraries
library(tidyverse)
library(here)
library(readxl)
library(janitor)
library(dataedu)
```

```{r}
# Read in using file path
# ExcelGradeBook <- read_excel("path/to/file.xlsx", sheet = 1, skip = 10)
```
# Read in the data
```{r}
# Use readxl package to read and import file and assign it a name
ExcelGradeBook <-
  read_excel(
    here("data", "gradebooks", "ExcelGradeBook.xlsx"),
    sheet = 1,
    skip = 10
  )
```

```{r}
# Rename data frame
gradebook <- ExcelGradeBook
```
# Begin data cleaning
# New functions
1. clean_names(), which takes messy column names that have periods, capitalized letters, spaces, etc. into R-friendly column names
2. get_dupes(), which identifies and examines duplicate records
3. tabyl(), which tabulates data in a data.frame format, and can be ‘adorned’ with the adorn_ functions to add total rows, percentages, and other dressings

```{r}
# look at the first few original column names
head(colnames(gradebook)) 
```

```{r}
# look at all the original column names
colnames(gradebook)
```

```{r}
# Clean the names with the clean_names() function
gradebook <- 
  gradebook %>% 
  clean_names()

# look at cleaned column names
head(colnames(gradebook))
```

```{r}
# Removing rows and columns with nothing but missing data
gradebook <- 
  gradebook %>% 
  remove_empty(c("rows", "cols"))

gradebook
```

```{r}
# Remove a targeted column because we don't use absent and late at this school.
gradebook <- 
  gradebook %>% 
  select(-absent, -late)

View(gradebook)
```
# Create new variables
We’ll use the function contains() from {stringr} to tell R to select columns that contain a certain string (that is, text). The function searches for any column with the string classwork_. The underscore makes sure the variables from classwork_1 all the way to classwork_17 are included in classwork_df.

```{r}
# Creates new data frame, selects desired variables from gradebook, and gathers all classwork scores into key/value pairs
classwork_df <-
  gradebook %>%
  select(
    name,
    running_average,
    letter_grade,
    homeworks,
    classworks,
    formative_assessments,
    projects,
    summative_assessments,
    contains("classwork_")) %>%
  # Note that scores are in character format. 
  # We use mutate() to transform them to numeric.
  mutate_at(vars(contains("classwork_")), list(~ as.numeric(.))) %>%
  pivot_longer(
    cols = contains("classwork_"),
    names_to = "classwork_number",
    values_to = "score"
  )
```

```{r}
# Summary of the data by columns
summary(gradebook)
```

```{r}
# Bar graph for categorical variable
gradebook %>%
  ggplot(aes(x = letter_grade,
             fill = running_average > 90)) +
  geom_bar() +
  labs(title = "Bar Graph of Student Grades",
       x = "Letter Grades",
       y = "Count",
       fill = "A or Better") +
  scale_fill_dataedu() +
  theme_dataedu()
```

